\chapter{Appendix I: Glossary}
\centering
\DefTblrTemplate{caption}{default}{}
\DefTblrTemplate{middlehead,lasthead}{default}{}
\begin{longtblr}[caption={Glossary},entry=none]{colspec={l|X}, rowhead=1}
    Term     & Definition                                                                                                         \\ \hline
    Agent    & An entity with agency (i.e. decision-making capability). Often used in terms of a simulation that individually models discrete states. \\
    BRT      & Bus Rapid Transit.
    A high frequency bus network with sections of exclusive right-of-way and priority at intersections.                           \\
    CSA      & Connection Scanning Algorithm. Timetabled pathfinding algorithm.                                                   \\
    Dijkstra & Dijkstra's pathfinding algorithm.                                                                                  \\
    DTP      & Department of Transport and Planning (Victorian Government Department).                                            \\
    GTFS     & General Transit Feed Specification. A standard for specifying public transport systems.                            \\
    MATSim   & Multi-Agent Transport Simulation. An extensible open-source simulator for agent-based microsimulations.            \\
    MC       & Multi-criteria (pathfinding problem extension).                                                                    \\
    RAPTOR   & Round bAsed Public Transit Optimised Router. Timetabled pathfinding algorithm.                                     \\
    TD       & Time-dependent.
    Used to refer to a graph representation of a timetabled network where each edge has a weight dependent on the departure time. \\
    TE       & Time-expanded.
    Used to refer to a graph representation of a timetabled network where each timetable stop is a separate node.                 \\
    VITM     & Victorian Integrated Transport Model.
    The comprehensive transport model used by the DTP to forecast travel demand across Victoria.                                  \\
\end{longtblr}
