\chapter{Glossary}
\begin{center}
\DefTblrTemplate{caption}{default}{}
\DefTblrTemplate{middlehead,lasthead}{default}{}
\begin{longtblr}[caption={Glossary},entry=none]{colspec={l|X}, rowhead=1}
    Term     & Definition                                                                                                         \\ \hline
    Agent    & An entity with agency (i.e. decision-making capability). Often used in terms of a simulation that individually tracks the states of many objects. \\
    Alighting & The act of getting on a vehicle. The opposite of "boarding".\\
    Connection & A particular instance of a vehicle moving from one stop to the next at a specific time.\\
    DTP      & Department of Transport and Planning (Victorian Government Department).                                            \\
    GTFS     & General Transit Feed Specification. A standard for specifying public transport systems.                            \\
    Link   & A pair of stations that are spatially adjacent.\\
    MC       & Multi-criteria (pathfinding problem extension). \\
    O-D Data & Origin-Destination data. A specific format of data, that encodes a list of journeys taken from the origin location to the destination location, at a particular departure time. \\
    RAPTOR   & Round bAsed Public Transit Optimised Router. The timetabled pathfinding algorithm implemented in WoB to assign agents to services.                                     \\
    WoB        & The "Who's on Board" novel rail transport model built in this project.
\end{longtblr}
\end{center}


%% Connections vs links 
% impute
% TrainSUM 